European or American? A re-examination of the transatlantic influence over the IMF

### 2025-07-30 manuscript/manuscript.qmd
europe lit review revised 2.0

However, while these studies offer important insights, they tend to remain case-specific or anecdotal. Crucially, they stop short of offering a systematic comparison between American and European influence, leaving open the question of how these two centres of power operate within the IMF.

>>>>

However, while these studies offer important insights, they tend to remain case-specific or anecdotal. While existing literature offers valuable case-based insights into either U.S. or European influence in IMF operations, there is a notable absence of systematic empirical comparisons between the two. This study addresses that gap by directly evaluating and contrasting how U.S. and European political ties affect IMF lending outcomes, including program access, conditionality, and loan volumes. Secondly, much of the literature treats the West as a monolithic bloc of creditor powers. However, this study disaggregates “the West” into the United States and Europe, revealing divergent motivations, instruments, and patterns of influence. In doing so, it challenges the assumption of homogeneous Western preferences and highlights variation in how and why influence is exerted.
